Stack Safety is a Security Property

### 2021-04-11 tex/main.tex
Passed over intro

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, leading to consequences
from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).} 
\fi
%
The protections offered by such mechanisms are generally described in terms
of a set of concrete examples of attacks that they can prevent, leaving a
more general characterization to the reader's intuition.
Moreover, the details of these mechanisms can be quite intricate
and often involve trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
Intuitively, memory cells within each stack frame are tagged with the
identity of
the function that owns the frame, and access to those locations is then gated
according to the function that is currently executing. Roessler and Dehon
explored a number of versions
of this basic idea, with different ways of identifying
functions and deciding exactly which accesses are allowed.  These alternatives
vary in their
efficiency and in the spatial and temporal precision of their security guarantees.
For example, their \emph{Depth Isolation} policy demands that a function
owns any
frame into which it writes; this seems simple and intuitively correct, but can be expensive to implement.
Conversely, their \emph{Lazy Tagging and Clearing} policy improves
performance by
allowing ``illicit'' writes to out-of-scope stack frames so long as the
written locations are not
subsequently read. The justification for this form of laziness is that any
genuinely dangerous writes will ultimately be
detected before they can affect program behavior. \bcp{Awkward transition here.} But we clearly need a
more formal foundation for
expressing, comparing and verifying these micro-policy security claims, as well as those of
other enforcement technologies.

@@

In this work, we take the (informal) definitions of WBCF and LSE as a starting point,
but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}.  We define
trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
In applying these security concepts, we treat each caller-callee pair as mutually distrustful
and view the caller as needing protection from the callee. \apt{Need more explanation of
  ``nesting'' and our novelty factor here? Also about domains??}
{\em Integrity} says that a caller's local data is protected from modification by its callee.
This is vital for local reasoning about the behavior of the caller.
{\em Confidentiality} states that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack.
Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
notion that data outside of live frames should be inaccessible (which LSE does not address).
We also give a simple trace-based formalization of WBCF, which is essential for keeping the
notions of ``caller' and ``callee'' sensible.

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces them.
\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
  significant contribution, and partly to give the reader some context for
  why this is part of our story.  E.g., ``One nice thing we can do with our
  spec is to use it to evaluate the claims of existing or future proposals
  for enforcement mechanisms.  One relatively inexpensive way to carry out
  such an evaluation is to see whether property-based testing against this
  spec can distinguish correct implementations of the mechanism from buggy
  ones...''}

However, their Lazy Tagging and Clearing micro-policy does not enforce stepwise integrity.
To address this, we develop an \emph{observational stack integrity} property that
is intended to specify enforcement mechanisms that
detect integrity violations between the point when they occur and the point when they make
an observable impact. Intuitively this property can be phrased as ``no visible reads after dangerous
writes.''  We prove (in Coq) that stepwise integrity implies observational integrity.

Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce this property either, as we show in
testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce observational integrity,
at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
pretty much done}

@@

Our initial description of the properties assumes single simple stack with no sharing between callers and
callees.  We expand it to allow passing of scalar stack data, granting a callee
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
Section 3, however it looks like things get tricky rather quickly.}

@@

While our property definitions were developed with micro-policy enforcement in mind, they
agnostic about many implementation details and should be applicable to other
systems with minor adjustment.\apt{move this up?}

@@

We begin with a description of our threat model
(\cref{sec:threat}), a running example that we refer to throughout the paper
(\cref{sec:running-example}),  and some technical preliminaries
(\cref{sec:prelim}).\apt{this para needs updating}

>>>>

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, with consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}
\fi
%
The implementation and configuration of these  stack-safety mechanisms can be complex,
and often involves compromises trading protection for efficiency.
Thus it would be useful to have a precise formal specification for stack
safety, both as a basis for comparing the security guarantees claimed by different enforcement
techniques and for verifying that these claims are met by particular implementations.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
In essence, memory locations within each stack frame are tagged with the identity of
the function activation that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. Many alternatives can
be implemented on top of this basic idea, by choosing how to identify activations
and how to maintain the connection between an activation identity and the memory it owns.
These alternatives vary in their
efficiency, but also in the spatial and temporal precision of their security guarantees.
For example, the \emph{Depth Isolation} policy requires a strict mapping from a function
to its frame from the start of the call to its end, and permits the function to write
into that frame. This seems simple and intuitively correct, but it can be expensive to
implement that strict mapping.
On the other hand, the \emph{Lazy Tagging and Clearing} policy improves performance by
assigning locations to the active function when they are written, which permits
illicit writes outside the function's frame so long as the proper owner of the written
locations does not subsequently read them. The intuition justifying laziness is that any
genuinely dangerous writes will ultimately be detected before they can affect program
behavior.  But we clearly need a more formal foundation for
expressing, comparing and verifying these micro-policy security claims, as well as those of
other enforcement technologies.

@@

In this work, we take the (informal) definitions of WBCF and LSE as a starting point,
but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}.
While our property definitions were developed with micro-policy enforcement in mind, they
agnostic about many implementation details and should be applicable to other
systems with minor adjustment. We define
trace-based properties of \emph{stepwise stack integrity} and \emph{stack confidentiality}.
In applying these security concepts, we treat each caller-callee pair as mutually distrustful
and view the caller as needing protection from the callee. \apt{Need more explanation of
  ``nesting'' and our novelty factor here? Also about domains??}
This differs from common instances in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{Thinking about how to express domains, or whether to. I'm not sure it's
  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
{\em Integrity} says that a caller's local data is protected from modification by its callee.
This is vital for local reasoning about the behavior of the caller.
{\em Confidentiality} states that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack.
Together, these properties (informally) capture LSE, as well as the ``temporal stack safety''
notion that data outside of live frames should be inaccessible (which LSE does not address).
We also give a simple trace-based formalization of WBCF, which is essential for keeping the
notions of ``caller'' and ``callee'' sensible.

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness.
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}

However, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
To address this, we develop an \emph{observational stack integrity} property that
is intended to specify enforcement mechanisms that
detect integrity violations between the point when they occur and the point when they make
an observable impact. Intuitively this property can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise integrity implies observational integrity.
\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
  established that this was important.}

Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce this property either, as we show in
testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}

@@

Our initial description of the properties assumes single simple stack with no sharing between
callers and callees.  We expand it to allow passing of scalar stack data, granting a callee
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
  Section 3, however it looks like things get tricky rather quickly.}
\sna{The trick is that if a programmer is passing pointers, we can't presume their
  intent -- how long should we permit memory to be shared? But if the compiler is passing
  pointers, as in pass by reference, we have a clear sense of where sharing should stop.}

@@

@@

We begin with a description of our threat model (\cref{sec:threat}),
then show how our properties apply to examples of stack unsafety 
(\cref{sec:running-example}). Then, after some technical preliminaries
(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}, \cref{sec:lazy})
and describe testing (\cref{sec:testing}).
